### 2016-12-11 paper/algorithms/cart.tex
paper: some cleanup, still missing motivating example and final experiments

  \begin{algorithmic}
    \Require{$T$ is a table. $D$ is a set of attributes of $T$ that need to be imputed, and
      $C$ is a set of attributes of $T$ that have complete data }
    \Ensure{Returns an imputed $T$}

    \Function{ImputeWithCE-CART}{$T, D, C$}
    	\State $T' \gets \Call{ImputeRandom}{T, D}$
	\For{$1 ... EPOCHS$}
		\For{$d \in D$}
            \State $imp \gets \Call{TrainRT}{\pi_{C \cup D \backslash \{d\}}(T'), \pi_d{T'}}$
            \State $T' \gets (\pi_{C \cup D \backslash \{d\}}(T'), \Call{PredictRT}{imp, T})$
		\EndFor
	\EndFor
	\Return{$T'$}
	\EndFunction
  \end{algorithmic}
  \caption{An algorithm for chained imputation using regression trees}
  \label{algo:imputation-strategy}