### 2017-02-07 paper/algorithms/plan-helpers.tex
paper: rewrite algorithm section for clarity

%\begin{algorithm}
  \begin{algorithmic}
    \Require{$q$ is a query plan. $C_{req}$ is the set of attributes that must be imputed in the output of this query plan. $C_{vis}$ is the set of attributes which are visible in the output of the overall query.}
    \Ensure{A set of query plans $Q$ with added imputation.}
    \Function{AddImpute}{$q, C_{req}, C_{vis}$}
    \State $D_{must} \gets \Call{Dirty}{q} \cap C_{req}$
    \State $D_{may} \gets D_{must} \cup (\Call{Dirty}{q} \cap C_{vis})$
    \State $Q \gets \emptyset$
    \If{$D_{must} = \emptyset$}
    \State $Q \gets Q \cup \{q\}$
    \Else
    \State $Q \gets Q \cup \{\mu_{D_{must}}(q),\ \delta_{D_{must}}(q)\}$
    \EndIf

    \If{$D_{may} \neq \emptyset$}
    \State $Q \gets Q \cup \{\mu_{D_{may}}(q)\}$
    \EndIf

    \State \Return $Q$
    \EndFunction

    \State

    \Require{$t$ is a table and $\phi$ is a filter predicate.}
    \Ensure{A set of plans for filtering $t$ with $\phi$.}
    \Function{OptFilter}{$t, \phi, C_{vis}$}
    \State \Return $\{\sigma_{\phi}(q)\ |\ q \in \Call{AddImpute}{t, \Call{Attrs}{\phi}, C_{vis}}\}$
    \EndFunction

    \State

    \Require{$q$ is a query plan, $G$ is a set of grouping attributes, $A$ is an aggregation function, and $C_{vis}$ is a set of visible attributes.}
    \Ensure{A set of plans for grouping \& aggregating with $G$ and $A$.}
    \Function{OptGroupBy}{$q, G, A, C_{vis}$}
    \State $C \gets G \cup \Call{Attrs}{A}$
    \State $Q' \gets \Call{AddImpute}{q, C, C_{vis}}$
    \State \Return $\{\Call{GroupBy}{q', G, A} ~|~ q' \in Q'\}$
    \EndFunction

    \State

    \Require{$q$ is a query plan, $P$ is a set of projection attributes.}
    \Ensure{A set of plans with the attributes in $P$ projected.}
    \Function{OptProject}{$q, P$}
    \State \Return $\{\pi_P(q') ~|~ q' \in \Call{AddImpute}{q, P, P}\}$
    \EndFunction

    \State

    \Require{A set of tables $T$, a plan cache $Q$ containing at least one query plan for each table in $T$, and a relation $J : T \times \Psi \times T$ between tables and join predicates.}
    \Ensure{An updated plan cache $Q'$ which contains at least one plan that combines all of the tables in $T$.}
    \Function{OptJoin}{$Q, T, J, C_{vis}$}
    \For{$size \in 2...|T|$}
    \For{$S \in \{\text{all length}\ size\ \text{subsets of}\ T\}$}
    \For{$t \in S$}
    \State $S' \gets S \setminus \{t\}$
    \For{$(t, \psi, t') \in J$ where $t' \in S'$}
    \State $C \gets \Call{Attrs}{\psi}$
    \State $L \gets \{\Call{AddImpute}{q, C, C_{vis}} ~|~ q \in Q[t] \}$
    \State $R \gets \{\Call{AddImpute}{q, C, C_{vis}} ~|~ q \in Q[S'] \}$
    \State $Q[S] \lhd \{l \bowtie_\psi r ~|~ l \in L, r \in R\}$
    \EndFor
    \EndFor
    \EndFor
    \EndFor
    \EndFunction
  \end{algorithmic}
  \caption{Base algorithms for query planning with imputations.}
  \label{algo:plan-helpers}
%\end{algorithm}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: