### 2017-02-22 paper/algorithms/plan-toplevel.tex
paper: rework algorithm and conclusion to account for new cost model

% \begin{algorithm}
\begin{algorithmic}[5]
  \Require{A set of tables $T$, a relation $F: T \times \Phi$ between tables and filter predicates, a relation $J : T \times \Psi \times T$ between tables and join predicates, a set of projection attributes $P$,  an optional set of grouping attributes $G$ and aggregator function $A$, and a parameter $\alpha \in [0, 1]$ that expresses trade-off performance and information loss}
  \Ensure{An optimized query plan.}

  \Function{Plan}{$T, F, J, P, G, A$}
  \State $C_g \gets \bigcup_{\phi \in F} \Call{Attr}{\phi}$ \Comment{Collect relevant attributes.} \label{lst:line:attr-start}
  \State $C_g \gets \bigcup_{\psi \in J} \Call{Attr}{\psi} \cup P \cup G \cup \Call{Attr}{A} \cup C_g$ \label{lst:line:attr-end}
  \State Let $Q$ be an empty plan cache.
  \For{$t \in T$} \Comment{Add selections to the plan cache.}
  \If {$\exists \phi: (t, \phi) \in F$}
  \State $Q[\{t\}] \lhd \Call{OptFilter}{Q, t, \phi, C_g}$
  \Else
  \State $Q[\{t\}] \lhd \{t\}$
  \EndIf
  \EndFor

  \State $\Call{OptJoin}{Q, T, J, C_g}$	\Comment{Optimize joins.}
  \State $B \gets Q[T]$ \Comment{Get the best plans for all tables.}
  \If{$G \neq \emptyset$} \Comment{Add optional group \& aggregate.}
  \State $B \gets \bigcup_{q \in B} \Call{OptGroupBy}{q, G, A, C_g}$
  \EndIf

  \State $B \gets \bigcup_{q \in B} \Call{OptProject}{q, P}$ \Comment{Add projections.}
  \State \Return $p \in B$ s.t. $p$ is $\alpha$-bound optimal  \Comment{Return fastest plan satisfying $\alpha$ bound.}
  \EndFunction

\end{algorithmic}
\caption{Top-level query planner with imputations.}
\label{algo:top-level-planner}
% \end{algorithm}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: