### 2017-05-11 paper/algorithms/histogram-transform.tex
paper: pass over cardinality estimation section

\newcommand{\InlineIte}[3]{#2\ \textbf{if}\ #1\ \textbf{else}\ #3}
\renewcommand{\algorithmicindent}{1em}

\begin{algorithmic}[1]
    \Require{
    \Statex
    \begin{itemize}[leftmargin=*,noitemsep]
    \item $H$: A map from attribute names to histograms.
    \item $op$: A relational operator.
    \end{itemize}}

  \Function{UpdateHistograms}{$H,\ op$}
  \If{op = $\delta_C$}\label{lst:drop-start}
  \State Set null count to zero for $H[c]$ where $c \in C$.
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]} = \min_d \Call{Card}{H[d]}$, $\forall c$.\label{lst:drop-end}
  \ElsIf{op = $\mu_C$}\label{lst:impute-start}
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]}$ is unchanged and null
  \State count is zero, $\forall c$.\label{lst:impute-end}
  \ElsIf{op = $\sigma_\phi$}\label{lst:std-start}
  \State Scale $H[c]$ by $\Call{Selectivity}{\phi, H}$, $\forall c$.
  \ElsIf{op = $\bowtie_\phi$}
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]} = \Call{Card}{\bowtie_\phi}$, $\forall c$.
  \EndIf\label{lst:std-end}
  \State \Return $H$
  \EndFunction
\end{algorithmic}
\caption{An algorithm for in-plan histogram updates}
\label{algo:histogram-transformation}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: