### 2017-05-15 paper/queries.tex
Add numbers to discussion of other imputation algs

\newcounter{queryno}
{
  \lstset{breaklines=true, basicstyle=\scriptsize\ttfamily}
\begin{tabular}{cl}
  \toprule
  \# & \multicolumn{1}{c}{Queries on CDC data} \\
  \midrule
  1 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(cuff_size) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 GROUP BY income;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q1} \\[-1.5ex]
  2 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines, basicstyle=\scriptsize\ttfamily]
SELECT income, AVG(creatine) FROM demo, exams, labs
WHERE demo.id = exams.id AND exams.id = labs.id
    AND income >= 13AND income <= 15 AND weight >= 63 GROUP BY income;
\end{lstlisting}
      \end{minipage}
      \refstepcounter{queryno}\label[query]{q2} \\[-1.5ex]
  3 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(blood_lead) FROM demo, exams, labs 
WHERE demo.id = labs.id AND labs.id = exams.id AND age_yrs <= 6;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q3}\\[-1.5ex]
  4 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT gender, AVG(blood_pressure_systolic) FROM demo, labs, exams 
WHERE demo.id = labs.id AND labs.id = exams.id AND body_mass_index >= 30 GROUP BY gender;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q4}\\[-1.5ex]
  5 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(waist_circumference) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 AND weight >= 100;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno}\label[query]{q5}\\
  \midrule
  \# & \multicolumn{1}{c}{Queries on freeCodeCamp data} \\
  \midrule
  6 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(income) FROM fcc WHERE income >= 50000 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q6} \\[-1.5ex]
  7 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(commutetime) FROM fcc WHERE gender = "female" AND countrycitizen = "United States";
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q7} \\[-1.5ex]
  8 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT schooldegree, AVG(studentdebtowe) FROM fcc
WHERE studentdebtowe > 0 and schooldegree >= 0 GROUP BY schooldegree;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q8}\\[-1.5ex]
  9 & 
      \begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(gdp_per_capita) FROM fcc, gdp 
WHERE fcc.countrycitizen = gdp.country AND age >= 18 GROUP BY attendedbootcamp;
\end{lstlisting}
      \end{minipage}\refstepcounter{queryno} \label[query]{q9}\\[-1.5ex]
  \bottomrule
\end{tabular}
}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: