### 2017-07-03 paper/algorithms/histogram-transform.tex
paper: rework algorithm 1

\newcommand{\InlineIte}[3]{#2\ \textbf{if}\ #1\ \textbf{else}\ #3}
\renewcommand{\algorithmicindent}{1em}

\review{
\cameraready{
\begin{algorithmic}[1]
    \Require{
    \Statex
    \begin{itemize}[leftmargin=*,noitemsep]
    \item $H$: A map from attribute names to histograms.
    \item $op$: A relational operator.
    \end{itemize}}

  \Function{UpdateHistograms}{$H,\ op$}
  \State Let $Cols$ be the set of attribute names in $H$.
  \If{op = $\delta_C$}\label{lst:drop-start}
  \State Set null count to zero for $H[c]$ for $c \in C$.
  \State $m \gets \min_{d \in \mathit{Cols}} \Call{Card}{H[d]}$
  \State Scale $H[d]$ s.t. $\Call{Card}{H[d]} = m$ for $d \in Cols$.\label{lst:drop-end}
  \ElsIf{op = $\mu_C$}\label{lst:impute-start}
  \For{$c \in C$}
  \State $k \gets \Call{Card}{H[c]}$
  \State Set null count of $H[c]$ to zero.
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]} = k$.
  \EndFor\label{lst:impute-end}
  \ElsIf{op = $\sigma_\phi$}\label{lst:std-start}
  \State Scale $H[d]$ by $\Call{Selectivity}{\phi, H}$ for $d \in Cols$.
  \ElsIf{op = $\bowtie_\phi$}
  \State Let $k_{join}$ be the estimated cardinality of the join.
  \State Scale $H[d]$ s.t. $\Call{Card}{H[d]} = k_{join}$ for $d \in Cols$.
  \EndIf\label{lst:std-end}
  \State \Return $H$
  \EndFunction
\end{algorithmic}
}
}
\caption{\review{An algorithm for in-plan histogram updates}}
\label{algo:histogram-transformation}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: